Exploratory Data Analysis of Kickstarter Data
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(lubridate)
```

# Description

I found this dataset on kaggle.com and thought it would be interesting to use for
Udacity's Exploratory Data Analysis project.

The set contains the first 4000 Kickstarter projects that were available on
Oct. 29, 2016. 


```{r echo=FALSE, Load_the_Data}
# Load the Data
ks <- read.csv(file="live.csv")
ks$end.time = as_date(ks$end.time)

start_time = week(as_date("2016-10-29"))
ks$weeks.until.end = week(ks$end.time) - start_time
ks$weeks.until.end <- factor(ks$weeks.until.end, levels = seq(8,0))

ggplot<- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Spectral")
def <- "#6699FF"
color <- "black"
```

# Univariate Plots Section


```{r echo=FALSE, Distribution_of_Percentage_Funded, warning=FALSE}
percent_funded_adjusted = quantile(ks$percentage.funded, 0.95)

ggplot(ks, aes(x = percentage.funded)) +
  geom_histogram(binwidth = 5, fill=def, color="black") +
  scale_x_continuous(breaks = seq(0, percent_funded_adjusted, 20), limits=c(0, percent_funded_adjusted)) +
  ylim(0, 400) + 
  labs(title="Distribution of Percentage Funded") +
  xlab("Percentage Funded (%)")

summary(ks$percentage.funded)
```
I first want to look at how many of our 4,000 projects have met or exceeded their
funding goals so I plotted a histogram of our percentage funded variable, a variable
that explains what percentage the project is funded in relation to its fundraising 
goals (ex: If I need $100 to print a cool sticker design and users have pledged $50, I am 50% funded)

As we can see, the mode occurs somewhere in our 0-5% bucket and the
median funded project is 18%. The average is much higher because we have extremely
skewed data due to some projects vastly exceeding 100% funding.

```{r echo=FALSE, Endtime, warning=FALSE}
ggplot(ks, aes(x=end.time)) +
  geom_histogram(bins=60, fill=def, color="black") +
  scale_x_date(date_breaks = "2 weeks") +
  labs(title="Distribution of End Times")
```

Of these 4,000 live projects, how long will they be available on Kickstarter? TO
answer this, I took a look at the end.time variable which has the date the project
will be removed from Kickstarter. It looks like at this given time, projects end
within the next month (this data was pulled on Oct. 29th). I wonder if the 
approaching holiday season had any influence here...

```{r echo=FALSE, Amount_Pledged_No_Outliers, warning=FALSE}
amt_pledged_adjusted = quantile(ks$amt.pledged, 0.90)

ggplot(ks, aes(x = amt.pledged)) +
  geom_histogram(binwidth = 100, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, amt_pledged_adjusted, 2000),limits = c(0, amt_pledged_adjusted)) + 
  ylim(0, 100) +
  labs(title="Distribution of Amount Pledged") +
  xlab("Amount Pledged ($)")

summary(ks$amt.pledged)
```

I was wondering how much funding these projects typically get so I plotted a histogram
of amounts pledged, the total amount of funding the project has at a given time. 
It looks like if you only need $1,000 to get your idea started, Kickstarter 
might be an appropriate platform.

```{r echo=FALSE, Amount_Pledged, warning=FALSE}
ggplot(ks, aes(x = amt.pledged)) +
  geom_histogram(binwidth = .1, fill=def, color="black") +
  scale_x_log10() + 
  labs(title="Distribution of Amount Pledged, with Log10 Scale")

summary(ks$amt.pledged)
```

Since the skew was so extreme on the last plot (I limited our x axis quite a bit)
I thought I would adjust the axis by log10 to get a better look at our distribution.

```{r echo=FALSE, warning=FALSE, Distribution_of_Title_Length}
ggplot(ks, aes(x = title.length)) +
  geom_histogram(binwidth = 1, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, max(ks$title.length), 5)) +
  labs(title="Distribution of Title Length") +
  xlab("Title Length in Characters")

summary(ks$title.length)
```
I wanted to see how long the average title was, is brevity key or do you want to 
use all space allotted to hook 'investors'? It looks like users typically use 
around 20 or 50 characters as our distribution is somewhat bimodal.

```{r echo=FALSE, warning=FALSE, Distribution_of_Blurb_Length}

ggplot(ks, aes(x = blurb.length)) +
  geom_histogram(binwidth = 1, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, max(ks$blurb.length), 10)) +
  labs(title="Distribution of Blurb Length") +
  xlab("Blurb length (# of Characters)")
```
While users may not use their entire title field, they are certainly taking 
advantage of the blurb! As we can see here, short blurbs are not very common.

```{r echo=FALSE, warning=FALSE, Distribution_Location_Type}
ggplot(ks, aes(x = type)) +
  geom_histogram(binwidth = 1, stat="count", fill=def, color="black") +
  scale_y_log10(breaks = 10^(0:3)) +
  labs(title="Distribution of Location Type") +
  xlab("Location Type")
```

The distibution of locations was strongly in favor of the 'Town' value, again I had
to use logarithmic scaling to represent the huge range of values.

```{r echo=FALSE, warning=FALSE, Distribution_Currency}
ggplot(ks, aes(x = currency)) +
  geom_histogram(binwidth = 1, stat="count", fill=def, color="black") +
  scale_y_log10(breaks = 10^(0:3)) +
  labs(title="Distribution of Currency Type")
```

We have a lot of currency option but it looks like Kickstarter is mostly popular
with Australia, US, CA, UK, and the EU. In the future, we will only want to use
these currencies.

```{r echo=FALSE, warning=FALSE, Country_Distribution}
top_countries <- table(ks$country)
top_countries <- as.data.frame(top_countries)
top_countries <- subset(top_countries, Freq > 3)

ggplot(subset(ks, country %in% top_countries$Var1), aes(x=country)) +
  geom_histogram(stat="count", fill=def, color="black") +
  scale_y_log10(breaks= 10^(0:3)) +
  labs(title="Distribution of Item Counts by Country")

```
It looks like our currency distributions match up with our items by country. Not
much Kickstarter usage in Asian countries.


# Univariate Analysis

### What is the structure of your dataset?

The dataframe is shaped 4000x13

Columns consist of the following types and values

```{r echo=FALSE, warning=FALSE, STR_OUT}
str(ks)
```

X is our index in the dataset
amt.pledged is the total amount of funding the project has at that time
blurb is a text field that describes the project being funded
by is the field containing the username of the user asking for funding
country is the user's country
currency is the user's native currency
end.time is the date the project will be closed for additional funding
location is string field for the users city, town, state, or really any location text
percentage.funded is the percentage describing amt pledged / amt needed 
state is the state or province the user is located in
title is the name of the project being funded
type is the location type with discrete values, town, island, county, etc
url is the link to the project on Kickstarter.com
title.length is a feature I engineered that has the length of the title in terms of character count.
blurb.length is another engineering feature that has the length of the blurb in terms of character count.
weeks.until.end is my final engineered feature that is the number of weeks until
  the project's funding period has ended. A value of 0 indicates this current week.


### What is/are the main feature(s) of interest in your dataset?

Amount pledged and Percent funded.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

I would like to see if there is any relationship between currency (as a substitue for location),
endtime, and our amount pledged and percent funded variables.

### Did you create any new variables from existing variables in the dataset?

I created title and blurb length columns as well as a weeks until end date column.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

* Distribution of Percentage Funded showed right skew
* I used a log10 scale when plotting Amount Pledged as there was very strong right skew
* Distribution of Title length is roughly bimodal with a mode at the title character limit, 60.
* Similar to title length, blurb length's mode appears to be at the character limit ~137.
  This plot has left skewness.
* Location type was predominately Town, currency was overwhelmingly USD, and item
country was also mostly US. All of these plots' y-axis have been adjust by log10.


# Bivariate Plots Section
```{r echo=FALSE, warning=FALSE, Title.length_vs_Blurb.length}
ggplot(ks, aes(x=title.length, y=blurb.length)) +
  geom_jitter(alpha=0.5) +
  scale_x_continuous(breaks=seq(0, 60, 10)) +
  scale_y_continuous(breaks=seq(0, 150, 20)) + 
  labs(title="Comparing Title and Blurb Lengths") +
  ylab("Blurb Length in Characters") +
  xlab("Title Length in Characters")
```

I wanted to visualize our text based columns relationship here, we already saw that
users are typically using the whole blurb field but not always the full title space,
here we can see that by the density of points at the top of the chart.

In the next few plots you will see me plot our text fields (blurb, title) character
lengths against our features of interest (amount pledged, percentage pledged). 
I was really hoping to see more of a relationship between the variables I created
and our features of interest but there does not appear to be much there.

```{r echo=FALSE, warning=FALSE, Title_Amt}
ggplot(ks, aes(x=title.length, y=amt.pledged)) +
  geom_jitter(alpha=0.5) +
  ylim(0, 450000) +
  scale_x_continuous(breaks=seq(0, 60, 10)) +
  labs(title="Comparing Title Lengths and Amt Pledged") +
  xlab("Title Length in Characters") +
  ylab("Amount Pledged in $")
```


```{r echo=FALSE, warning=FALSE, Title_Percent}
ggplot(ks, aes(x=title.length, y=percentage.funded)) +
  geom_jitter(alpha=0.5) +
  ylim(0, 1000) +
  scale_x_continuous(breaks=seq(0, 60, 10)) +
  labs(title="Comparing Title Lengths and Percentage Funded") +
  ylab("Percentage Funded %") +
  xlab("Title Length in Characters")
```

```{r echo=FALSE, warning=FALSE, Blurb_Amt}
ggplot(ks, aes(x=blurb.length, y=amt.pledged)) +
  geom_jitter(alpha=0.5) +
  ylim(0, 400000) +
  scale_x_continuous(breaks=seq(0, 150, 20)) +
  labs(title="Comparing Blurb Lengths and Amt Pledged") +
  xlab("Blurb Length in Characters") +
  ylab("Amount Pledged in $")
```

```{r echo=FALSE, warning=FALSE, Blurb_Percent}
ggplot(ks, aes(x=blurb.length, y=percentage.funded)) +
  geom_jitter(alpha=0.5) +
  ylim(0, 250) +
  scale_x_continuous(breaks=seq(0, 150, 20)) +
  labs(title="Comparing Blurb Lengths and Percentage Funded") +
  ylab("Percentage Funded %") +
  xlab("Blurb Length in Characters")
```


```{r echo=FALSE, Amt_Percent_Clean}
ggplot(ks, aes(x=amt.pledged, y=percentage.funded)) +
  geom_point(alpha=0.5) +
  coord_cartesian(xlim=c(0,50000), ylim=c(0, 250)) +
  labs(title="Comparing Amount Pledged vs Percent Funded") +
  xlab("Amount Pledged $") +
  ylab("Percentage Funded %")
```

Here we are looking at our 2 features of interest plotted against eachother.
It looks like most our projects under 50% funding receive less than $5,000, but
several projects need less than $5,000 to even meet their funding goals so we some
density following the y-axis up.

```{r echo=FALSE, warning=FALSE, PercentFunded_Endtime}
ggplot(ks, aes(x=end.time, y=percentage.funded)) +
  geom_jitter(alpha=0.5) +
  ylim(0, 250) +
  labs(title="Percentage Funded By End Time", x="Date") +
  ylab("Percentage Funded %")

```

In the plot above I wanted to see the trend between percentage funded and end time.
As we can see, projects with later end time have lower percentages of funding as
they likely have not been on Kickstarter as long as those with early November end
times.


```{r echo=FALSE, warning=FALSE, Pct_Complete_by_week}
ggplot(ks, aes(x=percentage.funded)) +
  geom_histogram(binwidth=5, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, 300, 50), limits = c(0, 300)) +
  ylim(0, 25) +
  facet_wrap(~weeks.until.end) +
  labs(title="Percentage Funded by Weeks Until End") +
  xlab("Percentage Funded %")
```

Here is that same distribution cut by week. In Week 1, it looks like we get a surge
of projects funded under 50%. I suspect these are likely joke products or quick 
fundraisers as there aren't nearly as many under funded projects in the previous
week.


```{r echo=FALSE, warning=FALSE, Amt_pledged_by_week}
ggplot(ks, aes(x=amt.pledged)) +
  geom_histogram(binwidth=2000, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, 100000, 25000), limits = c(0, 100000)) +
  ylim(0, 25) +
  facet_wrap(~weeks.until.end) +
  labs(title="Amt Pledged by Weeks Until End") +
  xlab("Amount Pledged $")
```


```{r echo=FALSE, warning=FALSE, Distribution_Pct_Funded}

interesting.currency <- subset(ks, ks$currency %in% c("aud", "cad", "usd", "eur","gdp"))

ggplot(interesting.currency, aes(x = amt.pledged)) +
  geom_histogram(binwidth = 1000, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, 25000, 5000), limits=c(0, 25000)) +
  facet_wrap(~currency, scales = "free_y")
```

It looks like most projects are receiving fewer than $5,000 in each of the Kickstarter
markets, with the US being the largest.

```{r echo=FALSE, warning=FALSE, Correlation}
cor(subset(ks, select = c(title.length, blurb.length, percentage.funded, amt.pledged)))
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

There is only one weak correlation in all of our data and that is between Amount
Pledged and Percentage Funded. Of course we would expect that the more money pledged
to your goal, the closer you are to reaching your funding goals.

Title and blurb length, the variables I created, did not prove to have much influence
on funding/percentage funded.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

Not particularly, I was hoping to find something that might serve as a good predictor
for reaching a value of 100 (or greater) in percentage funded, but I do not seem to 
have any yet.

I may have to revisit this and look for key words in titles and blurbs that might
have an association with higher percentage funded scores.

### What was the strongest relationship you found?

As you would expect it was a weak correlation between Amount Pledged and Percentage
Funded.


# Multivariate Plots Section

```{r echo=FALSE, warning=FALSE, Title_Blurb_Currency}
ggplot(interesting.currency, aes(x=title.length, y=blurb.length)) +
  geom_jitter(aes(color=currency)) +
  labs(title="Comparing Title and Blurb Lengths by Currency Used") +
  xlab("Title Length in Characters") + ylab("Blurb Length in Characters")
```
I was looking for clusters here to see if different currencies had a relationship
with blurb and title length. I.E. Are users in the European market more brief in their
blurb descriptions, etc. But there was nothing to be found aside from a reminder
that Kickstarter has a disproportionately large US user base.

```{r echo=FALSE, warning=FALSE, Amt_Per_currency}
ggplot(interesting.currency, aes(x=amt.pledged, y=percentage.funded)) +
  geom_point(aes(color=currency)) +
  coord_cartesian(xlim=c(0,50000), ylim=c(0, 250)) +
  labs(title="Comparing Amount Pledged and Percent Funded by Currency") +
  xlab("Amount Pledged $") +
  ylab("Percentage Funded %")
  
```
Now I am comparing currency with our main features to look for any relationships;
none to be found.

```{r echo=FALSE, warning=FALSE, Blurb_Per_Cur}
ggplot(interesting.currency, aes(x=blurb.length, y=percentage.funded)) +
  geom_point(aes(color=currency)) +
  scale_x_continuous(breaks=seq(0, 150, 25)) +
  ylim(0, 250) +
  labs(title="Comparing Blurb Length and Percent Funded by Currency") +
  xlab("Blurb Length in Characters") +
  ylab("Percentage Funded %")
```

There does not appear to be any relationship between currency and our Blurb vs
Percent Funded plot.

```{r echo=FALSE, warning=FALSE, Amt_Title_Currency}
ggplot(interesting.currency, aes(x=title.length, y=amt.pledged)) +
  geom_jitter(aes(color=currency)) +
  scale_y_continuous(breaks=seq(0,100000, 25000), limits = c(0, 100000)) +
  scale_x_continuous(breaks=seq(0, 60, 10)) +
  labs(title="Comparing Title Length and Percent Funded by Currency") +
  ylab("Amount Pledged in $") + xlab("Title Length in Words")
```

Again, looking for relationships with currency but I guess we can conclude that
users of differenty currency use Kickstarter similarly.

```{r echo=FALSE, warning=FALSE, Amt_pledge_week}
ggplot(ks, aes(x=amt.pledged, y=percentage.funded)) +
  geom_point(aes(color=weeks.until.end)) +
  coord_cartesian(xlim=c(0,25000), ylim=c(0, 300)) +
  labs(title="Comparing Amount Pledged and Percent Funded by Weeks Until End") +
  xlab("Amount Pledged $") +
  ylab("Percentage Funded %")
```
I like this plot, we can clearly see that as you move 'out' toward higher funded
projects, the end time is closer than our under funded projects. This is really
intuitive but it is neat to see visualized.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

I was looking for possible relationships between currency and our main features to 
see if there were any identifiable differences between regions and their investment
habits with KickStarter. Currency or country of origin has no discernable relationship
with either of our main features in this case.

### Were there any interesting or surprising interactions between features?

I wouldn't consider it surprising but in the Comparing Amount Pledged and Percent
Funded by Weeks Until Endtime, you can see that as you move toward higher values
along the x or y axis, points are generally darker, indicating that their funding
periods are coming to a close.

To dig further into this question I would have to dive a little deeper into the 
contents of the blurb or title and look for associations between keywords and amt funded or pledged.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, warning=FALSE, Plot_One}
ggplot(ks, aes(x = percentage.funded)) +
  geom_histogram(binwidth = 5, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, 300, 20), limits = c(0, 300)) +
  ylim(0, 400) + 
  labs(title="Distribution of Percentage Funded") +
  xlab("Percentage Funded %")

summary(ks$percentage.funded)
```

### Description One

Above is the distribution of our projects' percentage funded. Now, these are all
'live' projects meaning they are still open for contributions, but as we can see
in the end time distribution plot, most of these projects will end in the 4 weeks.

So, with that being said, we can infer that the median percentage funded for a project
on Kickstarter with a month left until end time, is 18% funded.


### Plot Two
```{r echo=FALSE, warning=FALSE, Plot_Two}
ggplot(ks, aes(x=percentage.funded)) +
  geom_histogram(binwidth = 10, fill=def, color="black") +
  scale_x_continuous(breaks=seq(0, 300, 50), limits = c(0, 300)) +
  ylim(0, 25) +
  facet_wrap(~ weeks.until.end) +
  labs(title="Percentage Funded by Weeks Until End") +
  xlab("Percentage Funded %")
```

### Description Two

In this plot I decided I would try to show percentage funded by the number of weeks 
left until the end time. We can clearly see that as weeks until end decreased, we 
saw a larger amount of our data shift to the right, indicating higher percentage 
of funding.

What I can't explain is why with 0 weeks left, we have a sudden surge of projects
that are funded below the 50% mark (whereas most projects with 1 week left are already
past the 50% mark). My guess is the user who created the posting was confused or
posted the item as a joke and chose an endtime < 7 days out.


### Plot Three
```{r echo=FALSE, warning=FALSE, Plot_Three}
ggplot(ks, aes(x=amt.pledged, y=percentage.funded)) +
  geom_point(aes(color=weeks.until.end))+
  coord_cartesian(xlim=c(0,25000), ylim=c(0, 300)) +
  labs(title="Comparing Amount Pledged and Percent Funded by Weeks Until End") +
  xlab("Amount Pledged $") +
  ylab("Percentage Funded %")
```

### Description Three

For plot 3 I wanted to visualize project funding in relation to how soon the project
ends.

I decided to use both percentage funded and amount pledged as variables describing
a well funded project and weeks.until.end as our measure for how soon the projects
are ending.

Since there is little correlation between any of our variables, this plot may be 
slightly confusing, but the point I am trying to convey is quite simple; the less
time a project has until it ends, the closer the project will be to meeting or 
exceeding its funding goals. The same rule applies to the amount pledged.

To reiterate, answer the following question: when would you expect a project to 
be more well funded; the day it is posted on Kickstarter or 3 weeks after it has
been posted?

The answer is almost in all cases 3 weeks after. The longer the project is available,
the more likely you are to receive pledges from the greater Kickstarter audience.

This is depicted in our visualization by any increase in the x or y axis. As you 
increase along or axis you will notice gradual color change from yellow to blue.

This is showing that as you look at points with more funding, or closer to their
funding goals, you typically find that they have fewer weeks left on Kickstarter.

It looks like projects under the 75% funding level are more likely to have 4 or more
weeks remaining until the project ends, while projects that exceed this percentage
funding level are often ending in fewer than 2 weeks.

------

# Reflection

Overall this was a fun project, I feel like I have some insight in the world of 
Kickstarter now.

### Successes and Failures

Successes, I was able to get a sense on
* the nationality of the Kickstarter userbase
* how long projects typically last
* % of projects that meet their funding goals
* Amounts pledged to a project

Failures, I was not able to
* Determine the types of products on kickstarter
* Find any strong correlation between our variables and meeting funding goals
* Find differences in user usage of Kickstarter by currency


### What we learned

* Kickstarter is a good source for funding if you need less than $4000
* Most listings take full advantage of title and blurb space
* Kickstarter is mostly popular in the US
* Median Percentage funded of items with less than 1 week until end time is 49%

### How we can improve

Something I left out of our exploration was that I sorted our percentage funded 
column descending and took a look at some of the most 'successful' products by 
this measure; most of these were either games or books of some variety.

I would like to be able to create a classifier that could examine the text contents in a 
title or blurb and determine the 'item type'. Possible item types from my quick
glance would be 'book', 'board game', 'card game', 'fashion accessory', 'technology
product', etc.